Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-01-26 CAV 2025 TODO/relatedwork.tex
related

Hybrid MILP can be seen as a refinement of $\alpha,\beta$-Crown~\cite{crown}, though its refined accurate path is vastly different than the base. This is not the case of Hybrid MILP, see Table \ref{table_hybrid}, which is much more accurate than $\alpha,\beta$-Crown.
That shortcoming for hard instances was witnessed in \cite{crown}, and a very specific solution using the full MILP encoding for the first few layers of a DNN was drafted, following similar proposal \cite{MILP2}. The main issue is that it is slow
and it cannot scale to DNNs with many neurons, as every neurons are encoded using an integer variable, making it not that accurate for intermediate networks (e.g. $9\times100$, $9\times200$, Table \ref{table_hybrid}), and not usable for larger DNNs ($6\times500$, CNN-B-Adv), whereas Hybrid MILP does scale.

Last, ERAN-DeepPoly \cite{deeppoly} computes bounds on values very quickly, by abstracting the weight of every node using two functions: an upper function and a lower function. While the upper function is fixed, the lower function offers two choices.
It relates to the LP encoding through the following new (to our knowledge) insight:  Proposition \ref{LP} state that the LP relaxation precisely matches the intersection of these two choices. Consequently, LP is more accurate (but slower) than DeepPoly, and Hybrid MILP is considerably more precise. Regarding PRIMA \cite{prima}, the approach involves explicitly maintaining dependencies between neurons.

@@

Finally, methods such as Reluplex / Marabou \cite{Reluplex,Marabou}  abstract the network: they diverge significantly from those abstracting values such as PRIMA, $\alpha,\beta$-CROWN)\cite{prima,crown}, Hybrid MILP. 
These network-abstraction algorithms are designed to be {\em complete} but completeness comes at the price of significant scalability challenges, and in practice they time-out on hard instances as shown in Table \ref{table_complete}.

>>>>

The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before \cite{MILP2}. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding, and then rely on a BaB call from that refined bounds (more complex calls to full MIPL and BaB). Non-surprisingly, this very slow technique does not scale but to small DNNs (max 2000 ReLU activation functions). Hybrid MILP on the other hand relies only on small calls: it is much more efficient on small DNNs, and it can scale to larger DNNs as well: we demonstrated strong performance with at least one order of magnitude larger networks (CNN-B-Adv).

%Hybrid MILP can be seen as a refinement of $\alpha,\beta$-Crown~\cite{crown}, though its refined accurate path is vastly different than the base. This is not the case of Hybrid MILP, see Table \ref{table_hybrid}, which is much more accurate than $\alpha,\beta$-Crown.
%That shortcoming for hard instances was witnessed in \cite{crown}, and a very specific solution using the full MILP encoding for the first few layers of a DNN was drafted, following similar proposal \cite{MILP2}. The main issue is that it is slow
%and it cannot scale to DNNs with many neurons, as every neurons are encoded using an integer variable, making it not that accurate for intermediate networks (e.g. $9\times100$, $9\times200$, Table \ref{table_hybrid}), and not usable for larger DNNs ($6\times500$, CNN-B-Adv), whereas Hybrid MILP does scale.

Last, ERAN-DeepPoly \cite{deeppoly} computes bounds on values very quickly, by abstracting the weight of every node using two functions: an upper function and a lower function. While the upper function is fixed, the lower function offers two choices.
It relates to the LP encoding through the following new (to our knowledge) insight:  Proposition \ref{LP} \cite{alessandro} states that the LP relaxation precisely matches the intersection of these two choices. Consequently, LP is more accurate (but slower) than DeepPoly, and Hybrid MILP is considerably more precise. Regarding PRIMA \cite{prima}, the approach involves explicitly maintaining dependencies between neurons.

@@

Finally, methods such as Reluplex / Marabou \cite{Reluplex,Marabou}  abstract the network: they diverge significantly from those abstracting values such as PRIMA, $\alpha,\beta$-CROWN)\cite{prima,crown}, Hybrid MILP. These network-abstraction algorithms are designed to be {\em intrinsically complete} (rather than asymptotically complete as BaB), but this comes at the price of significant scalability challenges, and in practice they time-out on complex instances as shown in Table \ref{table_complete}.
